A new tidy data structure to support exploration and modeling of temporal data

### 2019-01-29 main.Rmd
better animation

---
title: "A new tidy data structure to support exploration and modeling of temporal data"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Dianne
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob J
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types. On the other hand, the software supporting time series modeling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes. Going from raw data to model-ready data is painful. This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modeling and forecasting routines. Tidy data principles are extended to temporal data by: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time. This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based contexts. A sound data pipeline facilitates a fluent workflow for analyzing temporal data. The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
keywords: "time series, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: [references.bib, rpkgs.bib]
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

@@

# Software implementation {#sec:software}

@@

@@

```{r slide-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:slide-animate-cap)", interval = 0.3}
```

(ref:slide-animate-cap) An illustration of window of size 5 sliding over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)

```{r tile-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:tile-animate-cap)", interval = 0.3}
```

(ref:tile-animate-cap) An illustration of window of size 5 tiling over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)

```{r stretch-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:stretch-animate-cap)", interval = 0.3}
```

(ref:stretch-animate-cap) An illustration of window of size 5 stretching over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)

Figures \@ref(fig:slide-animate), \@ref(fig:tile-animate) and \@ref(fig:stretch-animate) show the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia. A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series.

>>>>

---
title: "A new tidy data structure to support exploration and modeling of temporal data"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Dianne
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob J
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types. On the other hand, the software supporting time series modeling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes. Going from raw data to model-ready data is painful. This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modeling and forecasting routines. Tidy data principles are extended to temporal data by: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time. This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based contexts. A sound data pipeline facilitates a fluent workflow for analyzing temporal data. The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
keywords: "time series, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: false
toc: false
bibliography: [references.bib, rpkgs.bib]
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

@@

# Software structure and design decisions {#sec:software}

@@

\autoref{fig:animate} shows the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia. A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series.

@@

```{r animate, fig.show = "animate", fig.height = 7, fig.cap = "(ref:animate-cap)", interval = 0.5, aniopts = c("autoplay", "controls", "loop")}
```

(ref:animate-cap) An illustration of window of size 5 rolling over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
